Variants on the Flight and Maintenance Planning problem

### 2018-07-16 latex/notes/complexite/FMP_variants.tex
modified the demonstration of the crew scheduling problem and un-commented the explanation of the two demonstration. It's still missing some parameters configuration and the objective function explanation for both.

@@

    For each period $k$ from the original problem, we create a period $t \in \mathcal{T}$. We can use the elements in these two sets indistinctly.
    For each period $k$ with demand $b_{k}$, we create a task $j$ with $T_j = \{t\}$ that requires $R_{j} = b_{t}| t \in T_j$ resources. This implies that $J_t = \{j\}| t \in T_j$, $|J_t|=1$ and $D_t = b_{t}$.

    For each worker $i$ we have a corresponding resource $i$.
    Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$.
    All resources start the planning horizon have just exited a maintenance in the previous period. This implies $N_t = 0$.
    The duration of the maintenance will be equal to $M = \overline{D}^{off} = \underline{D}^{off}$. This assumption is valid because the original proof was done with fixed values instead of ranges.

    The maximum remaining elapsed time after a maintenance is equal to $E = \overline{D}^{off}$. The minimum elapsed time before a maintenance is equal to $e=\underline{D}^{off}$.
    The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$. The minimum number of maintenances for each resource will be given by $M_{min}=\lceil(|\mathcal{T}| - \overline{D}^{work}) / M \rceil$. (Alternatively, the limit could be measured as a minimum number of maintenance months $M_{min}=|\mathcal{T}| - \overline{D}^{work}$).

@@

    \begin{tabular}{ll}
        $\mathcal{T}$ & Planning horizon periods. \\
        $\mathcal{I}$ & Resources. \\
        $\mathcal{J}$ & tasks.  \\
        $\mathcal{T}_j$ &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
    \end{tabular}

    \begin{tabular}{ll}
        $M_{min}$ & Minimum number of maintenances per resource in planning horizon. \\
        $M$ & Duration of maintenance. \\
        $E$ & Maximum elapsed time before assigning a maintenance. \\
        $E_{min}$ & Minimum elapsed time before assigning a maintenance. \\
        $R_{j}$ & Need of resources for task $j$. \\
    \end{tabular}

    % (here we could, in theory just copy the same model than we had in MOSIM).

        \begin{tabular}{ll}
            $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
            $ret_{it}$  &  remaining elapsed time (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
        \end{tabular}

        The following modifications need to be done to the original problem described in MOSIM

        \begin{align}
            & ret_{it} \in [E_{min},E] 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
            & \sum_{t \in \mathcal{T}} m_{it} \geq M_{min} 
                & i \in \mathcal{I} \label{eq:min_maintenace}
        \end{align}

        Constraints (\ref{eq:me}), which already existed, now have a specific minimum value instead of just 0. Constraints (\ref{eq:min_maintenace}) are new. They put a minimum number of maintenances for each resource. This had been considered as a possible cut to the original problem and does not imply a big change to the solutions.

        The alternative to constraints (\ref{eq:min_maintenace}), in case we want to control the months under maintenance and not the number of maintenance operations, are constraints (\ref{eq:min_maintenace2}).

        \begin{align}
            & \sum_{t \in \mathcal{T}} \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} \geq M_{min}
                & i \in \mathcal{I} \label{eq:min_maintenace2}\\
        \end{align}

        % \begin{align}
        %     % maximum capacity1:
        %     & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
        %       & t \in \mathcal{T} \label{eq:capacity1}\\
        %    %avail
        %    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
        %     &t \in \mathcal{T} \label{eq:avalaibility1}\\
        %    % maximum capacity2:                
        %     & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
        %             & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
        %     & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
        % \end{align}

        % \begin{align}
        %     % remaining used time
        %      & rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j \notag \\
        %             && t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
        %     & rut_{i0} = Rut^{Init}_i
        %            & i \in \mathcal{I} \label{eq:rut_initial}\\
        %     & rut_{it} \geq H m_{it}
        %             & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rut_lower}\\ 
        %     & rut_{it} \in [0,H]
        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} \\
        %     & ret_{it} \leq ret_{it-1} - 1 + E m_{it}
        %             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
        %     & ret_{i0} = Ret^{Init}_i
        %             & i \in \mathcal{I} \label{eq:ret_initial}\\
        %     & ret_{it} \geq E m_{it}
        %             & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
        %     & ret_{it} \in [0,E]
        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
        %     & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
        %           & t = |\mathcal{T}| \label{eq:min_ret}\\
        %     & \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
        %           & t = |\mathcal{T}| \label{eq:min_rut}
        % \end{align}
    % \begin{align}
    %     & \text{Min}\; \sum_{i \in \mathcal{I}} y_{i} \label{min_num_resources}
    % \end{align}

    % \begin{align}
    %     & \sum_{i \in \mathcal{I}} x_{it} \geq b_t & t \in \mathcal{T} \label{demand3} \\
    %     & \sum_{t \in \mathcal{T}} x_{it} \leq \overline{D}^{work} & i \in \mathcal{I} \label{max_work} \\
    %     & \sum_{t' = t}^{\overline{D}^{on}+t-1} x_{it'} \leq \overline{D}^{on} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_work_days} \\
    %     & \sum_{t' = t}^{\underline{D}^{on}+t-1} x_{it'} \geq \underline{D}^{on} (x_{it} - x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_work_days} \\
    %     & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \leq D^{off} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_no_work_days} \\
    %     & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \geq D^{off} (- x_{it} + x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_no_work_days} \\
    %     & y_{i} \geq x_{it} & i \in \mathcal{I}, t \in \mathcal{T} \label{num_resources} \\
    %     & x_{it}, y_{i} \in \{0, 1\} & i \in \mathcal{I}, t \in \mathcal{T} \label{binary}
    % \end{align}

    % Constraints (\ref{demand3}) guarantee the necessary number of workers is available to work. 
    % Constraints ((\ref{max_work}) checks that the total number of working days for each worker is less than the upper limit.
    % Constraints (\ref{max_work_days} - \ref{min_work_days}) control the resources meet the range of working days. Constraints (\ref{max_no_work_days} - \ref{min_no_work_days}) control the resources meet the fixed number of non-working days.
    % Finally constraints (\ref{num_resources}) quantify the number of necessary workers in the planning to be minimized in (\ref{min_num_resources}).

@@

    % For each employee $e \in E$, we create an analogous resource $i \in I$ which we will use indistinctly.

    % For each task $j \in J$, we create a $j \in J$ task in our problem with duration and minimal assignment time equal to $A^{min}_j = u_j$. Start times and finish times define the moment when the task is active: $\mathcal{T}_j = t \in \{s_j ... f_j\}$. The need of resources for each task will be constant $R_{j} = 1$. The number of resources working at any given time can $D_t = |\mathcal{J}_t|$ can be calculated. The relationship between tasks and employees is also equivalent: $\mathcal{O}_i = J_i$ and $\mathcal{I}_j = E_j$.

    % The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
    % For maintenances, we can just take them out of the equation completely. For this we just need to put: $E_{min} = E = M = N_t = 0$ and $Ret^{Init}_{i} = |\mathcal{T}| + 1$.

    % An instance of our problem is the following:

    % \subsection{Notations}

>>>>

    \begin{tabular}{ll}
        $\mathcal{E}$          & Employees. \\
        $\mathcal{K}$          & Planning horizon. \\
        $\overline{D}^{work}$  & Maximum number of working days at the of the planning horizon. \\
        $\underline{D}^{on}$   & Minimum number of consecutive working days for each employee. \\
        $\overline{D}^{on}$    & Maximum number of consecutive working days for each employee. \\
        $\underline{D}^{off}$  & Minimum number of consecutive non working days for each employee. \\
        $\overline{D}^{off}$   & Maximum number of consecutive non working days for each employee. \\
        $b_k$                  & Minimum number of workers to be assigned in period $k$. \\
    \end{tabular}

@@

        \begin{tabular}{l|l|l}
        \textbf{FMP}        &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
        $\mathcal{T}$       &  Planning horizon. & $ \mathcal{K} $ \\
        $\mathcal{I}$       &  Resources. & $ \mathcal{E} $ \\
        $\mathcal{J}$       &  Tasks.  & $ \{1 .. |\mathcal{K}|\} $ \\
        $\mathcal{T}_j$     &  time periods in which task $j$ is active. & $ t | t = j $ \\
        $\mathcal{J}_t$     &  tasks to be realized in period $t$. & $ j | j = t $ \\
        $\mathcal{I}_j$     &  resources that can be assigned to task $j$. & $ \mathcal{I} $\\
        $\mathcal{O}_i$     &  tasks for which resource $i$ can be used. & $ \mathcal{J} $ \\
        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ that control de maintenance starts. & $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ \\
    \end{tabular}

    \begin{tabular}{l|l|l}
        \textbf{FMP}      &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
        $D_t$             & number of resources to be assigned in total in period $t$ & $ b_k $ \\
        $A^{min}_j$       & minimum number of consecutive periods for task $j$. & $ 1 $ \\
        $R_{j}$           & Need of resources for task $j$ & $ b_k $ \\
        $H_j$             & amount of resource time required by task $j$. & $ 0 $ \\
        $M$               & maintenance duration in number of periods. & $ \underline{D}^{off} $ \\
        $E$               & remaining elapsed time after a maintenance. & $ \overline{D}^{on} $ \\
        $E^{min}$         & minimum elapsed time before a maintenance. & $ \underline{D}^{on} $ \\
        $H$               & remaining usage time after a maintenance. & $ 0 $ \\
        $N_t$             & number of resources in already-planned maintenances in period $t$ at the start. & $ 0 $ \\
        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start. & $ TODO $ \\
        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. & $ 0 $ \\
        $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. & $ 0 $\\
        $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. & $ 0 $ \\
    \end{tabular}

    % For each period $k$ from the original problem, we create a period $t \in \mathcal{T}$. We can use the elements in these two sets indistinctly.
    % For each period $k$ with demand $b_{k}$, we create a task $j$ with $T_j = \{t\}$ that requires $R_{j} = b_{t}| t \in T_j$ resources. This implies that $J_t = \{j\}| t \in T_j$, $|J_t|=1$ and $D_t = b_{t}$.

    % For each worker $i$ we have a corresponding resource $i$.
    % Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$.
    % All resources start the planning horizon have just exited a maintenance in the previous period. This implies $N_t = 0$.
    % The duration of the maintenance will be equal to $M = \overline{D}^{off} = \underline{D}^{off}$. This assumption is valid because the original proof was done with fixed values instead of ranges.

    % The maximum remaining elapsed time after a maintenance is equal to $E = \overline{D}^{off}$. The minimum elapsed time before a maintenance is equal to $e=\underline{D}^{off}$.
    % The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$. The minimum number of maintenances for each resource will be given by $M_{min}=\lceil(|\mathcal{T}| - \overline{D}^{work}) / M \rceil$. (Alternatively, the limit could be measured as a minimum number of maintenance months $M_{min}=|\mathcal{T}| - \overline{D}^{work}$).

@@

@@

    For each employee $e \in E$, we create an analogous resource $i \in I$ which we will use indistinctly.

    For each task $j \in J$, we create a $j \in J$ task in our problem with duration and minimal assignment time equal to $A^{min}_j = u_j$. Start times and finish times define the moment when the task is active: $\mathcal{T}_j = t \in \{s_j ... f_j\}$. The need of resources for each task will be constant $R_{j} = 1$. The number of resources working at any given time can $D_t = |\mathcal{J}_t|$ can be calculated. The relationship between tasks and employees is also equivalent: $\mathcal{O}_i = J_i$ and $\mathcal{I}_j = E_j$.

    The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
    For maintenances, we can just take them out of the equation completely. For this we just need to put: $E_{min} = E = M = N_t = 0$ and $Ret^{Init}_{i} = |\mathcal{T}| + 1$.
